### 2020-11-01 slides/slides/6_implementation(Cknown).tex
slides

\begin{frame}{Implementation (Easy case, known increment)}

%With \(n=3\) and \(\ell=20\), we prove that the Babai rounding solve exactly the CVP in our case.

%In practice, the Babai Rounding is still exact for \(\ell=19\)

  \begin{block}{Summary}
    \begin{itemize}
    \item Guess 37 bits:
      \begin{itemize}
      \item \(n=3\) successive rotations (6 bits each),
      \item \(\ell=19\) least significant bits of \(S_0\),
      \end{itemize}
    \item Solve \(2^{37}\) instances of CVP in dimension 3 (Babai Rounding).
    \item Reconstruct initial state, check.
    \end{itemize}
  \end{block}

  \begin{alertblock}{Caveat}
    Attack proved correct for $\ell=20$, works fine for $\ell=19$...
  \end{alertblock}

  \begin{exampleblock}{Concretely...}
    \begin{itemize}
    \item $25$ CPU cycles per guess, 23 CPU-minutes in total.
    \end{itemize}
  \end{exampleblock}

\end{frame}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: